---
title: "dvs audit log"
subtitle: "every dvs operation is appended to audit.log.jsonl in the storage directory"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

# Setup

```{r}
library(dvs)
library(fs)
library(here)
```

```{r}
source(here::here("R/mkdatasetfiles.R"))
```

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(storage)
dir.create(new_project)
```

Run a sequence of dvs operations to populate the log:

```{r}
#| warning: false
setwd(new_project)
mkdatasetfiles(n_files = 3, size_mb = 2, prefix = "file_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
dvs_init(storage)
```

```{r}
#| warning: false
setwd(new_project)
invisible(dvs_add("data/file_1.csv", message = "initial add of file 1"))
invisible(dvs_add("data/file_2.csv", message = "initial add of file 2"))
invisible(dvs_add("data/file_3.csv", message = "initial add of file 3"))
```

Simulate a restore (get):

```{r}
#| warning: false
setwd(new_project)
unlink("data/file_1.csv")
dvs_get("data/file_1.csv")
```

Modify `file_2.csv` and re-add:

```{r}
#| warning: false
setwd(new_project)
cat("99,0,0,0,0,0\n", file = "data/file_2.csv", append = TRUE)
invisible(dvs_add("data/file_2.csv", message = "updated file 2 — added row 99"))
```

# Reading the log

The audit log is a newline-delimited JSON file (`audit.log.jsonl`) in the
storage directory. Each line is one operation:

```{r}
audit_path <- fs::path(storage, "audit.log.jsonl")
```

Raw file — one JSON object per line:

```{r}
#| results: asis
lines <- readLines(audit_path)
cat("```json\n")
cat(lines, sep = "\n")
cat("\n```\n")
```

Same entries, pretty-printed:

```{r}
#| results: asis
cat("```json\n")
for (l in lines) {
  cat(jsonlite::toJSON(jsonlite::fromJSON(l, simplifyVector = FALSE), pretty = TRUE, auto_unbox = TRUE))
  cat("\n")
}
cat("```\n")
```

# Parsing the log

Each entry has the shape:

```json
{
  "operation_id": "<uuid>",
  "timestamp":    <unix-seconds>,
  "user":         "<username>",
  "action":       { "<command>": { ... } }
}
```

The `action` field is a tagged union — the single key is the command name
(`"init"`, `"add"`, or `"get"`). Parse it with `purrr` and `jsonlite`:

```{r}
#| warning: false
parse_entry <- function(line) {
  entry <- jsonlite::fromJSON(line)
  cmd   <- names(entry$action)[1]
  path  <- switch(cmd,
    add  = entry$action$add$file$path,
    get  = entry$action$get$file$path,
    NA_character_
  )
  list(
    operation_id = entry$operation_id,
    timestamp    = as.POSIXct(entry$timestamp, origin = "1970-01-01"),
    user         = entry$user,
    command      = cmd,
    path         = path
  )
}

audit_tbl <- readLines(audit_path) |>
  purrr::map(parse_entry) |>
  dplyr::bind_rows()

audit_tbl
```

# Querying the log

Operations by command type:

```{r}
dplyr::count(audit_tbl, command, sort = TRUE)
```

Files touched per operation (excluding `init`):

```{r}
#| warning: false
audit_tbl |>
  dplyr::filter(!is.na(path)) |>
  dplyr::select(timestamp, command, path) |>
  dplyr::arrange(timestamp)
```

How many times was each file touched?

```{r}
#| warning: false
audit_tbl |>
  dplyr::filter(!is.na(path)) |>
  dplyr::count(path, command, sort = TRUE)
```

# Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage,     recursive = TRUE)
```

---

**Next up**: [Random files](random_files.html) — the `mkdatasetfiles()` helper used to generate test data throughout these vignettes. Or return to the [full index](index.html).
