---
title: "dvs collaboration"
subtitle: "sharing data across projects and users through a common storage directory"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

# Overview

dvs separates *code* (tracked by git) from *data* (tracked by dvs). The
intended team workflow is:

1. **User A** adds data files → dvs copies blobs to **shared storage**, writes
   meta files under `.dvs/`
2. **User A** commits `dvs.toml` and the entire `.dvs/` tree to git (the meta
   files are small and text-based)
3. **User B** clones the repo. They get `dvs.toml` and `.dvs/` but *not* the
   data files (those are gitignored)
4. **User B** runs `dvs_get()` → dvs reads the meta files, locates blobs in
   shared storage, and writes the data files locally

The storage directory lives on shared infrastructure: a network mount,
an NFS/CIFS share, or any path your OS exposes as a local filesystem. In
this demo both users are simulated with separate temp dirs on the same
machine pointing at the same storage path.

# Setup

```{r}
options(width = 1000)
library(dvs)
library(fs)
library(here)
```

```{r}
source(here::here("R/mkdatasetfiles.R"))
```

```{r}
shared_storage <- basename(tempfile(fileext = "_storage"))
project_a      <- basename(tempfile(fileext = "_project_a"))
project_b      <- basename(tempfile(fileext = "_project_b"))
dir.create(here::here(shared_storage))
dir.create(here::here(project_a))
dir.create(here::here(project_b))
```

# User A: add data

```{r}
#| warning: false
setwd(here::here(project_a))
mkdatasetfiles(n_files = 3, size_mb = 1, prefix = "small_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
```

```{r}
#| warning: false
setwd(here::here(project_a))
dvs_init(here::here(shared_storage))
invisible(dvs_add(paths = fs::dir_ls("data", type = "file"), message = "add small corpus v1"))
dvs_status()
```

The `.dvs/` tree and `dvs.toml` are what user A would commit to git:

```{r}
#| warning: false
setwd(here::here(project_a))
fs::dir_tree(".dvs", all = TRUE, regexp = "\\.git", invert = TRUE)
```

```{r}
cat(readLines(here::here(project_a, "dvs.toml")), sep = "\n")
```

# Simulate a git clone (User B)

When user B clones the repo they receive `dvs.toml` and the `.dvs/` meta tree,
but `data/` is gitignored so the CSV files are absent:

```{r}
#| warning: false
fs::file_copy(here::here(project_a, "dvs.toml"), here::here(project_b, "dvs.toml"))
fs::dir_copy( here::here(project_a, ".dvs"),     here::here(project_b, ".dvs"))
```

User B's project shows all files as `absent`:

```{r}
#| warning: false
setwd(here::here(project_b))
dvs_status()
```

# User B: get data

```{r}
#| warning: false
setwd(here::here(project_b))
dvs_get(glob = "data/*.csv")
```

All files are now `current`:

```{r}
#| warning: false
setwd(here::here(project_b))
dvs_status()
```

Verify the content is byte-for-byte identical to what user A added:

```{r}
#| warning: false
files_a <- sort(fs::dir_ls(here::here(project_a, "data"), type = "file"))
files_b <- sort(fs::dir_ls(here::here(project_b, "data"), type = "file"))

identical(
  lapply(files_a, read.csv),
  lapply(files_b, read.csv)
)
```

# What belongs in git

| Path | Commit? | Notes |
|------|:-------:|-------|
| `dvs.toml` | ✅ | points to shared storage |
| `.dvs/**/*.dvs` | ✅ | meta files, the version ledger |
| `data/` | ❌ gitignore | large files live in dvs storage |
| `rv/library/` | ❌ gitignore | cached R packages rebuild from `rproject.toml` |
| `rv/scripts/` | ✅ | rv bootstrap helpers belong in the repo |

Minimal additions to `.gitignore`:

```text
data/
rv/library/
```

# Cleanup

```{r}
unlink(here::here(project_a),      recursive = TRUE)
unlink(here::here(project_b),      recursive = TRUE)
unlink(here::here(shared_storage), recursive = TRUE)
```

---

**Next up**: [Configuration](config.html): `dvs.toml` fields, compression tradeoffs, threads, and custom metadata folder names.
