---
title: "dvs configuration"
subtitle: "dvs.toml fields, compression tradeoffs, threads, and metadata folder name"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

# Setup

```{r}
options(width = 1000)
library(dvs)
library(fs)
library(here)
```

```{r}
source(here::here("R/mkdatasetfiles.R"))
```

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(storage)
dir.create(new_project)
```

```{r}
#| warning: false
setwd(new_project)
mkdatasetfiles(n_files = 5, size_mb = 5, prefix = "file_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
dvs_init(storage)
```

# `dvs.toml`

`dvs_init()` writes a `dvs.toml` at the project root:

```{r}
cat(readLines(fs::path(new_project, "dvs.toml")), sep = "\n")
```

| Field | Meaning |
|-------|---------|
| `storage_path` | Where dvs copies blobs. Any directory reachable from this machine — local path or network mount. |
| `metadata_folder_name` | Defaults to `.dvs`. The folder that mirrors your data tree and holds meta files. Change if `.dvs` collides with another tool. |
| `compression` | `"zstd"` (default) or `"none"`. zstd typically halves storage footprint for tabular data. |

# Compression: `zstd` vs `none`

Same data, two different compression settings — compare resulting storage size.

```{r}
storage_zstd <- tempfile(fileext = "_zstd_storage", tmpdir = here::here())
storage_none <- tempfile(fileext = "_none_storage", tmpdir = here::here())
project_zstd <- tempfile(fileext = "_zstd_project", tmpdir = here::here())
project_none <- tempfile(fileext = "_none_project", tmpdir = here::here())
for (d in c(storage_zstd, storage_none, project_zstd, project_none)) dir.create(d)
```

```{r}
#| warning: false
setwd(project_zstd)
mkdatasetfiles(n_files = 5, size_mb = 10, prefix = "data_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
```

```{r}
#| warning: false
setwd(project_none)
mkdatasetfiles(n_files = 5, size_mb = 10, prefix = "data_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
```

```{r}
#| warning: false
setwd(project_zstd)
dvs_init(storage_zstd, compression = "zstd")
invisible(dvs_add(glob = "data/*.csv", message = "add with zstd compression"))
```

```{r}
#| warning: false
setwd(project_none)
dvs_init(storage_none, compression = "none")
invisible(dvs_add(glob = "data/*.csv", message = "add with no compression"))
```

Compare storage sizes:

```{r}
#| warning: false
storage_summary <- function(path, label) {
  info <- fs::dir_info(path, recurse = TRUE, type = "file") |>
    dplyr::filter(!grepl("\\.git|audit\\.log", path))
  dplyr::tibble(
    compression  = label,
    n_blobs      = nrow(info),
    size_bytes   = sum(as.numeric(info$size)),
    size_mb      = round(sum(as.numeric(info$size)) / 1e6, 2)
  )
}

comparison <- dplyr::bind_rows(
  storage_summary(storage_zstd, "zstd"),
  storage_summary(storage_none, "none")
)

comparison |>
  dplyr::mutate(ratio = round(size_bytes / size_bytes[compression == "none"], 3))
```

# Threads

`set_dvs_threads()` controls how many parallel threads dvs uses for hashing
and copying blobs. The default is auto-detect (one thread per logical CPU).

```{r}
storage_t <- tempfile(fileext = "_threads_storage", tmpdir = here::here())
project_t <- tempfile(fileext = "_threads_project", tmpdir = here::here())
dir.create(storage_t)
dir.create(project_t)
```

```{r}
#| warning: false
setwd(project_t)
mkdatasetfiles(n_files = 50, size_mb = 3, prefix = "f_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
dvs_init(storage_t)
```

Time `dvs_add` with a single thread:

```{r}
#| warning: false
set_dvs_threads(1)
t1 <- system.time({
  setwd(project_t)
  invisible(dvs_add(glob = "data/*.csv", message = "1 thread"))
})
cat("1 thread :", round(t1["elapsed"], 2), "s\n")
```

Re-stage files by appending a byte so dvs sees them as unsynced:

```{r}
#| warning: false
setwd(project_t)
for (f in fs::dir_ls("data", type = "file")) write("", f, append = TRUE)
```

Time with four threads:

```{r}
#| warning: false
set_dvs_threads(4)
t4 <- system.time({
  setwd(project_t)
  invisible(dvs_add(glob = "data/*.csv", message = "4 threads"))
})
cat("4 threads:", round(t4["elapsed"], 2), "s\n")
```

# Custom metadata folder name

`metadata_folder_name` overrides the default `.dvs` folder. Useful when
another tool in the project already claims that name.

```{r}
storage_m <- tempfile(fileext = "_meta_storage", tmpdir = here::here())
project_m <- tempfile(fileext = "_meta_project", tmpdir = here::here())
dir.create(storage_m)
dir.create(project_m)
```

```{r}
#| warning: false
setwd(project_m)
mkdatasetfiles(n_files = 1, size_mb = 1, prefix = "d_", dir = "data", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
dvs_init(storage_m, metadata_folder_name = "datalock")
invisible(dvs_add(glob = "data/*.csv", message = "custom folder name demo"))
```

The meta files live under `datalock/` instead of `.dvs/`:

```{r}
setwd(project_m)
fs::dir_tree("datalock", all = TRUE)
```

The `dvs.toml` records the custom name so all subsequent operations pick it up
automatically:

```{r}
cat(readLines(fs::path(project_m, "dvs.toml")), sep = "\n")
```

# `group` (Unix only)

`dvs_init(storage, group = "mygroup")` sets the Unix group on the storage
directory and all blobs written to it. This ensures that members of the group
can read and write blobs without permission errors on shared filesystems.

```r
# example — not executed here
dvs_init("/mnt/shared/storage", group = "data-team")
```

# Cleanup

```{r}
for (d in c(storage, new_project,
            storage_zstd, storage_none, project_zstd, project_none,
            storage_t, project_t,
            storage_m, project_m)) {
  unlink(d, recursive = TRUE)
}
```

---

**Next up**: [Audit log](audit.html) — parsing and querying `audit.log.jsonl` with purrr and dplyr.
