---
title: "dvs Error Reference"
subtitle: "every error dvs can raise, shown side-by-side for R and the CLI"
format: html
execute:
  freeze: auto
---

# Setup

```{r}
library(dvs)
library(here)
```

Create a temporary storage directory and project root, then expose them to the
shell so the bash chunks below can use `$DVS_PROJECT` and `$DVS_STORAGE`.

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())

dir.create(storage)
dir.create(new_project)

# Make it a git repo so dvs can find the project root
system2("git", c("init", new_project), stdout = FALSE, stderr = FALSE)

Sys.setenv(DVS_PROJECT = new_project, DVS_STORAGE = storage)
```

# `init` errors

## Storage path is inside the repository

Passing a storage path that resolves to a subdirectory of the project root is
rejected — dvs refuses to version-control its own blob store.

```{r}
#| error: true
setwd(new_project)
dvs_init(storage_path = file.path(new_project, "inside_storage"),
         root_dir     = new_project)
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs init "$DVS_PROJECT/inside_storage"
```

## Project already initialised — `dvs.toml` exists

Once `dvs.toml` is present a second `init` call fails immediately.

```{r}
#| error: true
# First init succeeds
setwd(new_project)
dvs_init(storage_path = storage, root_dir = new_project)

# Second init fails
dvs_init(storage_path = storage, root_dir = new_project)
```

```{bash}
#| error: true
# dvs.toml already created by the R chunk above
cd "$DVS_PROJECT"
dvs init "$DVS_STORAGE"
```

## Errors not demonstrated here

- **Backend storage already exists** (`dvs is already initialized (backend storage exists)`): triggered by removing `dvs.toml` and `.dvs/` while leaving the storage directory intact, then re-running init. Requires filesystem surgery; straightforward to contrive but not illuminating.

---

# `add` errors

## Not inside a DVS project

Calling `dvs_add` / `dvs add` from a directory that has no `dvs.toml` anywhere
in its ancestor chain fails before touching the filesystem.

```{r}
#| error: true
# Use a fresh temp dir with no dvs.toml
outside <- tempfile(tmpdir = here::here())
dir.create(outside)
setwd(outside)

dvs_add(paths = "anything.csv")
```

```{bash}
#| error: true
OUTSIDE=$(mktemp -d)
cd "$OUTSIDE"
dvs add anything.csv
rm -rf "$OUTSIDE"
```

## No files to add — empty glob

A glob pattern that matches no files returns an empty set; dvs raises an error
rather than silently doing nothing.

```{r}
#| error: true
setwd(new_project)
dvs_add(glob = "data/*.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs add --glob "data/*.csv"
```

## File does not exist

Requesting a specific path that is not present on disk is a per-file error;
the overall command exits non-zero.

```{r}
#| error: true
setwd(new_project)
dvs_add(paths = "nonexistent_file.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs add nonexistent_file.csv
```

## Path resolves to a directory, not a file

Passing a plain directory name (without a glob) is rejected.

```{r}
#| error: true
setwd(new_project)
dir.create(file.path(new_project, "mydir"), showWarnings = FALSE)
dvs_add(paths = "mydir")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
mkdir -p mydir
dvs add mydir
```

## Path is outside the project root

An explicit path that canonicalises to a location above (or beside) the project
root is rejected with a per-file error.

```{r}
#| error: true
setwd(new_project)
outside_file <- tempfile(fileext = ".csv")
writeLines("a,b,c", outside_file)
# Use a relative path that escapes the project root
rel_outside <- file.path("..", basename(outside_file))
dvs_add(paths = rel_outside)
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
OUTSIDE_FILE=$(mktemp --suffix=.csv)
echo "a,b,c" > "$OUTSIDE_FILE"
dvs add "$OUTSIDE_FILE"
```

## Errors not demonstrated here

- **Path not found during glob resolution** (`Path not found: <path>`): fired by `resolve_paths_for_add` when an explicit path argument does not exist on disk at all (distinct from a valid-but-untracked path). The "file does not exist" demo above exercises the post-resolution `file not found` variant; the pre-resolution path is harder to distinguish at the R / shell surface level.
- **Failed to resolve path** (symlink / IO error): requires a broken symlink or unavailable mount; environment-specific.

---

# `status` errors

## Not inside a DVS project

`dvs_status` / `dvs status` require a `dvs.toml` in the ancestor tree.

```{r}
#| error: true
outside2 <- tempfile(tmpdir = here::here())
dir.create(outside2)
setwd(outside2)
dvs_status()
```

```{bash}
#| error: true
OUTSIDE2=$(mktemp -d)
cd "$OUTSIDE2"
dvs status
rm -rf "$OUTSIDE2"
```

## Errors not demonstrated here

- **Unparseable meta file** (`error: failed to determine relative path` / serde JSON error): requires manually corrupting a `.dvs` metadata file. The error surfaces as a per-file `StatusDetail::Error` row rather than a top-level abort.

---

# `get` errors

## Not inside a DVS project

```{r}
#| error: true
outside3 <- tempfile(tmpdir = here::here())
dir.create(outside3)
setwd(outside3)
dvs_get(paths = "something.csv")
```

```{bash}
#| error: true
OUTSIDE3=$(mktemp -d)
cd "$OUTSIDE3"
dvs get something.csv
rm -rf "$OUTSIDE3"
```

## No files to get — empty glob

When the glob resolves to zero tracked files, `dvs get` aborts.

```{r}
#| error: true
setwd(new_project)
# No files have been added yet under data/; glob matches nothing in metadata
dvs_get(glob = "data/nope_*.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs get --glob "data/nope_*.csv"
```

## File is not tracked by DVS

Requesting a file that exists on disk but has never been `dvs add`-ed.

```{r}
#| error: true
setwd(new_project)
writeLines("col1,col2", file.path(new_project, "untracked.csv"))
dvs_get(paths = "untracked.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
echo "col1,col2" > untracked.csv
dvs get untracked.csv
```

## File is not found at all

Requesting a path that neither exists on disk nor has metadata.

```{r}
#| error: true
setwd(new_project)
dvs_get(paths = "ghost.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs get ghost.csv
```

## Storage blob missing for tracked file

The metadata `.dvs` file exists but the corresponding blob has been deleted
from storage — simulating a corrupted or partially-synced store.

```{r}
#| error: true
setwd(new_project)

# Add a real file so we have metadata
writeLines("x,y,z", file.path(new_project, "tracked.csv"))
dvs_add(paths = "tracked.csv")

# Remove the local copy so get actually attempts a retrieve
file.remove(file.path(new_project, "tracked.csv"))

# Delete every blob from storage so the retrieve has nothing to copy
blob_dirs <- list.dirs(storage, recursive = TRUE, full.names = TRUE)
blob_files <- list.files(storage, recursive = TRUE, full.names = TRUE)
blob_files <- blob_files[!grepl("audit", blob_files)]
invisible(file.remove(blob_files))

# Now get should fail: metadata exists but blob is gone
dvs_get(paths = "tracked.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
# tracked.csv was deleted and its blob removed by the R chunk above
dvs get tracked.csv
```

## Errors not demonstrated here

- **Retrieved file does not match expected hash** (`Retrieved file does not match expected hash`): requires writing a blob with corrupted content under the correct hash path, so decompression succeeds but the post-retrieve hash check fails. Needs direct filesystem manipulation of the content-addressed store.

---

# Cleanup

```{r}
setwd(here::here())
unlink(new_project, recursive = TRUE)
unlink(storage,     recursive = TRUE)
```
