---
title: "dvs Error Reference"
subtitle: "every error dvs can raise, shown side-by-side for R and the CLI"
format: html
execute:
  freeze: auto
---

# Setup

```{r}
library(dvs)
library(here)

# Print tibbles wide so all columns appear on one row.
# options(width = Inf) is rejected by base R (max 10000); 10000 is the ceiling.
options(width = 10000)
```

Create a temporary storage directory and project root, then expose them to the
shell so the bash chunks below can use `$DVS_PROJECT` and `$DVS_STORAGE`.

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())

dir.create(storage)
dir.create(new_project)

# Make it a git repo so dvs can find the project root
system2("git", c("init", new_project), stdout = FALSE, stderr = FALSE)

Sys.setenv(DVS_PROJECT = new_project, DVS_STORAGE = storage)
```

# `init` errors

## Storage path is inside the repository

Passing a storage path that resolves to a subdirectory of the project root is
rejected — dvs refuses to version-control its own blob store.

```{r}
#| error: true
setwd(new_project)
dvs_init(storage_path = file.path(new_project, "inside_storage"),
         root_dir     = new_project)
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs init "$DVS_PROJECT/inside_storage"
```

## Project already initialised — `dvs.toml` exists

Once `dvs.toml` is present a second `init` call fails immediately.

```{r}
#| error: true
# First init succeeds
setwd(new_project)
dvs_init(storage_path = storage, root_dir = new_project)

# Second init fails
dvs_init(storage_path = storage, root_dir = new_project)
```

```{bash}
#| error: true
# dvs.toml already created by the R chunk above
cd "$DVS_PROJECT"
dvs init "$DVS_STORAGE"
```

## Backend storage already exists

If `dvs.toml` is deleted but the storage directory still contains a previous
backend layout, re-running init refuses rather than silently reusing the
backend.

```{r}
#| error: true
# Use fresh temp dirs so we don't disrupt the main demo project
bse_storage <- tempfile(fileext = "_storage", tmpdir = here::here())
bse_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(bse_storage); dir.create(bse_project)
system2("git", c("init", bse_project), stdout = FALSE, stderr = FALSE)

setwd(bse_project)
dvs_init(storage_path = bse_storage, root_dir = bse_project)

# Remove dvs.toml but leave the storage intact
file.remove(file.path(bse_project, "dvs.toml"))

# Second init detects the existing backend
dvs_init(storage_path = bse_storage, root_dir = bse_project)
```

```{bash}
#| error: true
BSE_STORAGE=$(mktemp -d)
BSE_PROJECT=$(mktemp -d)
git -C "$BSE_PROJECT" init -q
cd "$BSE_PROJECT"
dvs init "$BSE_STORAGE" > /dev/null
rm dvs.toml
dvs init "$BSE_STORAGE"
rm -rf "$BSE_STORAGE" "$BSE_PROJECT"
```

# `add` errors

## Not inside a DVS project

Calling `dvs_add` / `dvs add` from a directory that has no `dvs.toml` anywhere
in its ancestor chain fails before touching the filesystem.

```{r}
#| error: true
outside <- tempfile(tmpdir = here::here())
dir.create(outside)
setwd(outside)

dvs_add(paths = "anything.csv")
```

```{bash}
#| error: true
OUTSIDE=$(mktemp -d)
cd "$OUTSIDE"
dvs add anything.csv
rm -rf "$OUTSIDE"
```

## No files to add — empty glob

A glob pattern that matches no files raises in the R package; the CLI prints
`No files to add` and exits 0.

```{r}
#| error: true
setwd(new_project)
dvs_add(glob = "data/*.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs add --glob "data/*.csv"
```

## Path not found

Passing an explicit path that doesn't exist on disk aborts before any files
are processed — neither surface does partial work.

```{r}
#| error: true
setwd(new_project)
dvs_add(paths = "nonexistent_file.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs add nonexistent_file.csv
```

## Path resolves to a directory (no glob)

Passing a bare directory name with no glob matches no files. The R package
raises, the CLI prints `No files to add` and exits 0 — mild surface
divergence.

```{r}
#| error: true
setwd(new_project)
dir.create(file.path(new_project, "mydir"), showWarnings = FALSE)
dvs_add(paths = "mydir")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
mkdir -p mydir
dvs add mydir
```

## Path is outside the project root

You're inside the project but ask dvs to track a file that lives elsewhere on
disk, by absolute path. R returns a 2-column tibble — `path` + `error` — with
the failure recorded as a row rather than an R-level condition.

```{r}
#| error: true
setwd(new_project)
outside_file <- tempfile(fileext = ".csv")
writeLines("a,b,c", outside_file)
res <- dvs_add(paths = outside_file)
res |> print(width = Inf, n = Inf)
```

Inspect the result columns directly — `path` is the offending path, `error`
is a plain character string:

```{r}
str(res)
res$path
res$error
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
OUTSIDE_FILE=$(mktemp -t dvs_outside).csv
echo "a,b,c" > "$OUTSIDE_FILE"
dvs add "$OUTSIDE_FILE"
rm -f "$OUTSIDE_FILE"
```

## Mixed paths: one exists, one does not

When only some paths exist, both surfaces abort at the first missing path —
no partial add.

```{r}
#| error: true
setwd(new_project)
writeLines("x", file.path(new_project, "present.csv"))
dvs_add(paths = c("present.csv", "missing.csv"))
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
echo "x" > present.csv
dvs add present.csv missing.csv
```

# `status` errors

## Not inside a DVS project

`dvs_status` / `dvs status` require a `dvs.toml` in the ancestor tree.

```{r}
#| error: true
outside2 <- tempfile(tmpdir = here::here())
dir.create(outside2)
setwd(outside2)
dvs_status()
```

```{bash}
#| error: true
OUTSIDE2=$(mktemp -d)
cd "$OUTSIDE2"
dvs status
rm -rf "$OUTSIDE2"
```

## Path outside the project root

Unlike `add`, an absolute path outside the project does **not** produce an
error for `status` — the path just doesn't match anything tracked and you
get the empty-set response.

```{r}
#| error: true
setwd(new_project)
outside_status <- tempfile(fileext = ".csv")
writeLines("a,b,c", outside_status)
dvs_status(paths = outside_status) |> print(width = Inf, n = Inf)
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
OUTSIDE_S=$(mktemp -t dvs_outside_s).csv
echo "a,b,c" > "$OUTSIDE_S"
dvs status "$OUTSIDE_S"
rm -f "$OUTSIDE_S"
```

## Mixed paths: one tracked, one does not exist

Unlike `add`, `status` quietly drops paths it can't match — you get a tibble
with only the tracked path and no error or warning about the missing one.

```{r}
#| error: true
setwd(new_project)
# Track one real file so we have something to match against
writeLines("x", file.path(new_project, "present.csv"))
dvs_add(paths = "present.csv")

dvs_status(paths = c("present.csv", "missing.csv")) |>
  print(width = Inf, n = Inf)
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs status present.csv missing.csv
```

## Unparseable meta file

If a `.dvs` metadata file is corrupted (non-JSON content), the parse error
surfaces as a per-file row in the result rather than a top-level abort — the
rest of the tree still resolves.

```{r}
#| error: true
# Fresh project so we can corrupt one meta file without affecting the rest
ump_storage <- tempfile(fileext = "_storage", tmpdir = here::here())
ump_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(ump_storage); dir.create(ump_project)
system2("git", c("init", ump_project), stdout = FALSE, stderr = FALSE)

setwd(ump_project)
dvs_init(storage_path = ump_storage, root_dir = ump_project)
writeLines("x", file.path(ump_project, "corrupt.csv"))
dvs_add(paths = "corrupt.csv")

# Corrupt the meta file
writeLines("garbage", file.path(ump_project, ".dvs", "corrupt.csv.dvs"))

dvs_status() |> print(width = Inf, n = Inf)

Sys.setenv(UMP_PROJECT = ump_project)
```

```{bash}
#| error: true
cd "$UMP_PROJECT"
dvs status
```

# `get` errors

## Not inside a DVS project

```{r}
#| error: true
outside3 <- tempfile(tmpdir = here::here())
dir.create(outside3)
setwd(outside3)
dvs_get(paths = "something.csv")
```

```{bash}
#| error: true
OUTSIDE3=$(mktemp -d)
cd "$OUTSIDE3"
dvs get something.csv
rm -rf "$OUTSIDE3"
```

## No files to get — empty glob

When the glob resolves to zero tracked files, `dvs get` aborts.

```{r}
#| error: true
setwd(new_project)
dvs_get(glob = "data/nope_*.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs get --glob "data/nope_*.csv"
```

## File is not tracked by DVS

Requesting a file that exists on disk but has never been `dvs add`-ed.

```{r}
#| error: true
setwd(new_project)
writeLines("col1,col2", file.path(new_project, "untracked.csv"))
dvs_get(paths = "untracked.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
echo "col1,col2" > untracked.csv
dvs get untracked.csv
```

## File is not found at all

Requesting a path that neither exists on disk nor has metadata.

```{r}
#| error: true
setwd(new_project)
dvs_get(paths = "ghost.csv")
```

```{bash}
#| error: true
cd "$DVS_PROJECT"
dvs get ghost.csv
```

## Storage blob missing for tracked file

The metadata `.dvs` file exists but the corresponding blob has been deleted
from storage — simulating a corrupted or partially-synced store.

```{r}
#| error: true
# Use a fresh project; the main demo project still has live state we don't
# want to wreck for downstream chunks.
bm_storage <- tempfile(fileext = "_storage", tmpdir = here::here())
bm_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(bm_storage); dir.create(bm_project)
system2("git", c("init", bm_project), stdout = FALSE, stderr = FALSE)

setwd(bm_project)
dvs_init(storage_path = bm_storage, root_dir = bm_project)
writeLines("x,y,z", file.path(bm_project, "tracked.csv"))
dvs_add(paths = "tracked.csv")

# Remove the local copy and delete every blob from storage
file.remove(file.path(bm_project, "tracked.csv"))
blob_files <- list.files(bm_storage, recursive = TRUE, full.names = TRUE)
invisible(file.remove(blob_files[!grepl("audit", blob_files)]))

dvs_get(paths = "tracked.csv") |> print(width = Inf, n = Inf)

Sys.setenv(BM_PROJECT = bm_project)
```

```{bash}
#| error: true
cd "$BM_PROJECT"
dvs get tracked.csv
```

## Retrieved file does not match expected hash

The blob exists and decompresses successfully, but its content hashes to
something other than the hash recorded in the meta file. Simulates
content-level corruption of the blob.

```{r}
#| error: true
# Fresh project + storage
hm_storage <- tempfile(fileext = "_storage", tmpdir = here::here())
hm_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(hm_storage); dir.create(hm_project)
system2("git", c("init", hm_project), stdout = FALSE, stderr = FALSE)

setwd(hm_project)
dvs_init(storage_path = hm_storage, root_dir = hm_project)
writeLines("original", file.path(hm_project, "tampered.csv"))
dvs_add(paths = "tampered.csv")

# Replace the stored blob with zstd-compressed different content under the
# same filename (so the hash-keyed path resolves, but the decompressed bytes
# no longer match the recorded hash).
blob <- list.files(hm_storage, recursive = TRUE, full.names = TRUE)
blob <- blob[!grepl("audit", blob)]
Sys.chmod(blob, mode = "0644")
# zstd -q reads stdin and writes compressed bytes to stdout
system2("bash", c("-c",
                  sprintf("printf 'DIFFERENT BYTES' | zstd -q > %s",
                          shQuote(blob))))

# Remove local file so get actually goes to storage
file.remove(file.path(hm_project, "tampered.csv"))

dvs_get(paths = "tampered.csv") |> print(width = Inf, n = Inf)

Sys.setenv(HM_PROJECT = hm_project)
```

```{bash}
#| error: true
cd "$HM_PROJECT"
dvs get tampered.csv
```

# Cleanup

```{r}
setwd(here::here())
unlink(new_project, recursive = TRUE)
unlink(storage,     recursive = TRUE)
for (d in c("bse_project", "bse_storage", "ump_project", "ump_storage",
            "bm_project", "bm_storage", "hm_project", "hm_storage")) {
  if (exists(d)) unlink(get(d), recursive = TRUE)
}
```
