---
title: "Getting Started with dvs CLI"
subtitle: "init, add, status, and get: five minutes end to end"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

## Setup

R creates the sandbox directories and dataset; all dvs operations below run via
the `dvs` CLI binary.

```{r}
options(width = 1000)
library(here)
```

Create isolated temporary directories for this demo:

```{r}
storage     <- basename(tempfile(fileext = "_storage"))
new_project <- basename(tempfile(fileext = "_project"))
dir.create(here::here(storage))
dir.create(here::here(new_project))
```

Save the built-in `Theoph` pharmacokinetic dataset as a CSV:

```{r}
dir.create(here::here(new_project, "data"))
write.csv(Theoph, here::here(new_project, "data/theoph.csv"), row.names = FALSE)
```

Expose the paths to the shell:

```{r}
Sys.setenv(DVS_PROJECT = here::here(new_project), DVS_STORAGE = here::here(storage))
```

## `dvs init`

Initialize a dvs repository pointing at the storage directory:

```{bash}
cd "$DVS_PROJECT"
dvs init "$DVS_STORAGE"
cat dvs.toml
```

## `dvs status` (before any adds)

No files are tracked yet:

```{bash}
cd "$DVS_PROJECT"
dvs status
```

## `dvs add`

Track the dataset with a message:

```{bash}
cd "$DVS_PROJECT"
dvs add data/theoph.csv -m "add Theoph dataset"
```

## `dvs status`

The file is now tracked and `current`:

```{bash}
cd "$DVS_PROJECT"
dvs status
```

## Add a second copy

Write another copy of the dataset and track it:

```{r}
write.csv(Theoph, here::here(new_project, "data/theoph_v2.csv"), row.names = FALSE)
```

```{bash}
cd "$DVS_PROJECT"
dvs add data/theoph_v2.csv -m "add second Theoph copy"
dvs status
```

## Modify a file, check status

Append a row to `theoph.csv` to simulate a local edit after it was stored:

```{bash}
echo "99,70.5,320,0,0.1" >> "$DVS_PROJECT/data/theoph.csv"
```

`dvs status` detects that the file on disk no longer matches the stored hash:

```{bash}
cd "$DVS_PROJECT"
dvs status
```

## Delete the original, check status

```{bash}
cd "$DVS_PROJECT"
rm data/theoph.csv
dvs status
```

`theoph.csv` shows `absent`; `theoph_v2.csv` remains `current`.

## `dvs get`

Restore the deleted file from storage:

```{bash}
cd "$DVS_PROJECT"
dvs get data/theoph.csv
```

Both files are `current` again:

```{bash}
cd "$DVS_PROJECT"
dvs status
```

## Cleanup

```{r}
unlink(here::here(new_project), recursive = TRUE)
unlink(here::here(storage),     recursive = TRUE)
```

---

**Next up**: [Getting Started, R API](getting-started.html): the same workflow using the `dvs` R package functions directly.
