---
title: "R walkthrough"
subtitle: "The core loop with library(dvs)"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

This walkthrough runs the everyday loop from R: initialize a project, add a
file, check its status, delete it, then get it back. It is the "I re-cloned a
project and need its data" sequence. The [CLI walkthrough](getting-started-cli.html)
covers the same steps from the terminal.

## Setup

```{r}
options(width = 1000)
library(dvs)
library(here)
```

Create a project directory and a sibling storage directory, then save R's
built-in `Theoph` dataset as a CSV.

```{r}
storage     <- basename(tempfile(fileext = "_storage"))
new_project <- basename(tempfile(fileext = "_project"))
dir.create(here::here(storage))
dir.create(here::here(new_project))
dir.create(here::here(new_project, "data"))
write.csv(Theoph, here::here(new_project, "data/theoph.csv"), row.names = FALSE)
```

## 1. Initialize

Point storage at the sibling directory. See [dvs_init()](r-init.html) for every
parameter.

```{r}
setwd(here::here(new_project))
dvs_init(here::here(storage))
```

## 2. Add

Track the CSV with a message. See [dvs_add()](r-add.html).

```{r}
setwd(here::here(new_project))
dvs_add("data/theoph.csv", message = "initial Theoph data")
```

## 3. Status

The file is tracked and `current`. See [dvs_status()](r-status.html).

```{r}
setwd(here::here(new_project))
dvs_status()
```

## 4. Remove the file

Delete the local copy, the moment after a fresh clone where the data is not on
disk yet.

```{r}
unlink(here::here(new_project, "data/theoph.csv"))
```

## 5. Status again

The file is now `absent`: tracked, but missing locally.

```{r}
setwd(here::here(new_project))
dvs_status()
```

::: {.callout-note}
dvs reports three states: `current`, `absent`, and `unsynced` (on disk but not
matching the stored hash). See [Storage and meta files](intro-internals.html).
:::

## 6. Get

Restore the file from storage. See [dvs_get()](r-get.html).

```{r}
setwd(here::here(new_project))
dvs_get("data/theoph.csv")
```

## 7. Status again

The file is `current` again. The loop is complete.

```{r}
setwd(here::here(new_project))
dvs_status()
```

## Cleanup

```{r}
unlink(here::here(new_project), recursive = TRUE)
unlink(here::here(storage),     recursive = TRUE)
```

## Next steps

- The [R Package](r-init.html) reference covers every parameter of
  [dvs_init()](r-init.html), [dvs_add()](r-add.html),
  [dvs_status()](r-status.html), and [dvs_get()](r-get.html).
- The same loop from the terminal: [CLI walkthrough](getting-started-cli.html).
- How dvs stores data: [Storage and meta files](intro-internals.html).
