---
title: "Getting Started with dvs"
subtitle: "init, add, status, and get — five minutes end to end"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

## Setup

```{r}
options(width = 1000)
library(dvs)
library(fs)
library(here)
```

Create isolated temporary directories for this demo:

```{r}
storage     <- basename(tempfile(fileext = "_storage"))
new_project <- basename(tempfile(fileext = "_project"))
dir.create(here::here(storage))
dir.create(here::here(new_project))
```

Save the built-in `Theoph` pharmacokinetic dataset as a CSV:

```{r}
dir.create(here::here(new_project, "data"))
write.csv(Theoph, here::here(new_project, "data/theoph.csv"), row.names = FALSE)
```

## `dvs_init`

Initialize a dvs repository pointing at the storage directory:

```{r}
setwd(here::here(new_project))
dvs_init(here::here(storage))
```

## `dvs_status` (before any adds)

No files are tracked yet:

```{r}
setwd(here::here(new_project))
dvs_status()
```

## `dvs_add`

Track the dataset with a message:

```{r}
setwd(here::here(new_project))
dvs_add("data/theoph.csv", message = "add Theoph dataset")
```

## `dvs_status`

The file is now tracked and `current`:

```{r}
setwd(here::here(new_project))
dvs_status()
```

## Add a second copy

Write another copy of the dataset and track it:

```{r}
write.csv(Theoph, here::here(new_project, "data/theoph_v2.csv"), row.names = FALSE)
```

```{r}
setwd(here::here(new_project))
dvs_add("data/theoph_v2.csv", message = "add second Theoph copy")
dvs_status()
```

## Modify a file, check status

Append a row to `theoph.csv` to simulate a local edit after it was stored:

```{r}
write.csv(
  rbind(Theoph, Theoph[1L, ]),
  here::here(new_project, "data/theoph.csv"),
  row.names = FALSE
)
```

`dvs_status` detects that the file on disk no longer matches the stored hash:

```{r}
setwd(here::here(new_project))
dvs_status()
```

## Delete the original, check status

```{r}
unlink(here::here(new_project, "data/theoph.csv"))
```

`theoph.csv` now shows `absent`; `theoph_v2.csv` remains `current`:

```{r}
setwd(here::here(new_project))
dvs_status()
```

## `dvs_get`

Restore the deleted file from storage:

```{r}
setwd(here::here(new_project))
dvs_get("data/theoph.csv")
```

Both files are `current` again:

```{r}
setwd(here::here(new_project))
dvs_status()
```

## Cleanup

```{r}
unlink(here::here(new_project), recursive = TRUE)
unlink(here::here(storage),     recursive = TRUE)
```

---

**Next up**: [Getting Started — CLI](getting-started-cli.html) — the same workflow driven from the terminal via the `dvs` binary.
