---
title: "Storage and meta files"
subtitle: "blob layout, hashing, meta JSON, and compression"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

::: {.callout-note}
This page is implementation detail, beyond what normal use requires. You do not
need any of it to run the four verbs.
:::

This page builds a project with files of several sizes, then inspects the
metadata folder and the storage directory to show how dvs lays data out: a
content-addressed blob store, a `.dvs` meta file per tracked file, and per-file
compression.

# Setup

```{r}
options(width = 1000)
library(dvs)
library(fs)
library(here)
```


# Helpers

```{r}
source(here::here("R/mkdatasetfiles.R"))
```


```{r}
storage     <- basename(tempfile(fileext = "_storage"))
new_project <- basename(tempfile(fileext = "_project"))
dir.create(here::here(storage))
dir.create(here::here(new_project))
```

```{r}
#| warning: false
setwd(here::here(new_project))

mkdatasetfiles(n_files = 1,  size_mb = 25, prefix = "large_",      dir = "data/large",      show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
mkdatasetfiles(n_files = 3,  size_mb = 1,  prefix = "small_",      dir = "data/small",      show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
mkdatasetfiles(n_files = 50, size_mb = 3,  prefix = "individual_", dir = "data/individual", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH")))
```

```{r}
#| warning: false
setwd(here::here(new_project))

dvs_init(here::here(storage), compression = "none")

dvs_add("data/large/large_1.csv", message = "add one large dataset")
dvs_add(paths = fs::dir_ls("data/small", type = "file"), message = "add small corpus datasets")
dvs_add(glob = "data/individual/individual_*.csv", message = "add individual datasets via glob")
```

# Internals of `.dvs`

`.dvs` is hidden, so `fs::` helpers need `all = TRUE`.

```{r}
#| warning: false
setwd(here::here(new_project))

fs::dir_tree(".dvs", all = TRUE, regexp = "\\.git", invert = TRUE)
```

```{r}
#| warning: false
setwd(here::here(new_project))

fs::dir_info(".dvs", recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::select(size, path, type)
```

```{r}
#| warning: false
setwd(here::here(new_project))

fs::dir_info(".dvs", recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::mutate(subdir = path |> fs::path_dir() |> fs::path_rel(".dvs")) |>
  dplyr::reframe(n_files = dplyr::n(), size = sum(size), .by = subdir)
```

# Meta files

Each tracked data file has a `.dvs` meta file under `.dvs/`, mirroring the data tree.

```{r}
#| warning: false
setwd(here::here(new_project))

meta_files <- fs::dir_ls(".dvs", recurse = TRUE, all = TRUE, glob = "*.dvs", type = "file")
meta_files <- meta_files[!grepl("\\.git", meta_files)]
meta_files
```

Contents of the meta file for `data/large/large_1.csv`:

```{r}
#| warning: false
setwd(here::here(new_project))

large_meta <- grep("large_1", meta_files, value = TRUE)
large_meta
```

```{r}
#| results: asis
#| warning: false
setwd(here::here(new_project))
cat("```json\n")
cat(jsonlite::toJSON(jsonlite::fromJSON(large_meta, simplifyVector = FALSE), pretty = TRUE, auto_unbox = TRUE))
cat("\n```\n")
```
# Storage directory

```{r}
fs::dir_tree(here::here(storage), all = TRUE, regexp = "\\.git", invert = TRUE)
```

```{r}
fs::dir_info(here::here(storage), recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::mutate(subdir = path |> fs::path_dir() |> fs::path_rel(here::here(storage))) |>
  dplyr::reframe(n_files = dplyr::n(), size = sum(size), .by = subdir)
```

# Hashed blob is plain text

Pick up the blake3 hash from the meta file.

```{r}
#| warning: false
setwd(here::here(new_project))

meta <- jsonlite::fromJSON(large_meta)
meta

hash <- meta$hashes$blake3
hash
```

Blobs are split into `<first-2>/<rest>`:

```{r}
#| warning: false

large_blob <- here::here(storage, substr(hash, 1, 2), substr(hash, 3, nchar(hash)))
large_blob
fs::file_exists(large_blob)
```

Initialized with `compression = "none"`, the stored object is the original CSV, addressable by hash. Readable as-is.

```{r}
#| warning: false

readLines(large_blob, n = 5)
```

```{r}
readLines(
  n = 5, here::here(new_project, "data", "large", "large_1.csv")
)
```



# `.dvs/` mirrors `data/`

Meta files live under `.dvs/` at the same relative paths as the data files they track.

```{r}
#| warning: false
setwd(here::here(new_project))

fs::dir_tree("data", regexp = "\\.git", invert = TRUE)
```

```{r}
#| warning: false
setwd(here::here(new_project))

fs::dir_tree(".dvs", all = TRUE, regexp = "\\.git", invert = TRUE)
```

# Cleanup

```{r}
unlink(here::here(new_project), recursive = TRUE)
unlink(here::here(storage),     recursive = TRUE)
```

## See also

- [The dvs.toml project file](config.html): the project configuration.
- [The audit log](audit.html): the append-only record of every add.
- The command references: [dvs_add()](r-add.html) / [dvs add](cli-add.html).
