---
title: "Introduction to dvs CLI"
subtitle: "the same operations driven from the terminal via the `dvs` binary"
format: html
execute:
  freeze: auto
---

# Setup

R creates the sandbox directories and data files; all dvs operations below run
via the `dvs` CLI binary.

```{r}
library(dvs)
library(fs)
library(here)
```

```{r}
source(here::here("R/mkdatasetfiles.R"))
```

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(storage)
dir.create(new_project)
```

```{r}
#| warning: false
setwd(new_project)
mkdatasetfiles(n_files = 1,  size_mb = 25, prefix = "large_",      dir = "data/large")
mkdatasetfiles(n_files = 3,  size_mb = 1,  prefix = "small_",      dir = "data/small")
mkdatasetfiles(n_files = 50, size_mb = 3,  prefix = "individual_", dir = "data/individual")
```

Pass the paths to the shell so bash chunks can use them:

```{r}
Sys.setenv(DVS_PROJECT = new_project, DVS_STORAGE = storage)
```

# `dvs init`

```{bash}
cd "$DVS_PROJECT"
dvs init "$DVS_STORAGE"
cat dvs.toml
```

# `dvs add`

Add a single file with a message:

```{bash}
cd "$DVS_PROJECT"
dvs add data/large/large_1.csv -m "add one large dataset"
```

Add a directory of files via glob:

```{bash}
cd "$DVS_PROJECT"
dvs add --glob "data/small/*.csv" -m "add small corpus datasets"
```

Add 50 files at once:

```{bash}
cd "$DVS_PROJECT"
dvs add --glob "data/individual/individual_*.csv" -m "add individual datasets"
```

## `--dry-run`

Preview what *would* be added without writing anything:

```{bash}
cd "$DVS_PROJECT"
dvs add --glob "data/small/*.csv" --dry-run
```

## Meta file

Each tracked file gets a `.dvs` meta file that records its hash:

```{bash}
cd "$DVS_PROJECT"
cat .dvs/data/large/large_1.csv.dvs
```

# `dvs status`

Default — show all tracked files:

```{bash}
cd "$DVS_PROJECT"
dvs status
```

Filter by status:

```{bash}
cd "$DVS_PROJECT"
dvs status --current
```

Show extended metadata columns:

```{bash}
cd "$DVS_PROJECT"
dvs status --with-metadata
```

Delete a few files to produce `absent` entries:

```{bash}
cd "$DVS_PROJECT"
rm data/individual/individual_04.csv data/individual/individual_08.csv data/individual/individual_15.csv
dvs status --absent
```

Modify a file to produce an `unsynced` entry:

```{bash}
cd "$DVS_PROJECT"
echo "99,0,0,0,0,0" >> data/large/large_1.csv
dvs status --unsynced
```

## JSON output

Every command accepts `--json` for machine-readable output:

```{bash}
cd "$DVS_PROJECT"
dvs status --absent --json
```

# `dvs get`

Restore absent files by glob:

```{bash}
cd "$DVS_PROJECT"
dvs get --glob "data/individual/individual_*.csv"
dvs status --absent
```

## `--dry-run`

```{bash}
cd "$DVS_PROJECT"
rm data/individual/individual_01.csv
dvs get --glob "data/individual/individual_*.csv" --dry-run
```

# `--threads`

Override the auto-detected thread count for any command:

```{bash}
cd "$DVS_PROJECT"
dvs status --threads 2
```

# Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage,     recursive = TRUE)
```

---

**Next up**: [Internals](intro-internals.html) — see how dvs organises meta files and blob storage under the hood.
