---
title: "Introduction to dvs internals"
format: html
execute:
  freeze: auto
---

# Data project

Same project as `vignette/intro.qmd`: 1x25MB `large`, 3x1MB `small`, 50x3MB `individual` CSVs under `data/`, with a dvs repository initialized against an external `storage` directory.

# Setup

```{r}
library(dvs)
library(fs)
library(here)
```


# Helpers

```{r}
source(here::here("R/mkdatasetfiles.R"))
```


```{r}
storage <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(storage)
dir.create(new_project)
```

```{r}
#| warning: false
setwd(new_project)

mkdatasetfiles(n_files = 1,  size_mb = 25, prefix = "large_",      dir = "data/large")
mkdatasetfiles(n_files = 3,  size_mb = 1,  prefix = "small_",      dir = "data/small")
mkdatasetfiles(n_files = 50, size_mb = 3,  prefix = "individual_", dir = "data/individual")
```

```{r}
#| warning: false
setwd(new_project)

dvs_init(storage, compression = "none")

dvs_add("data/large/large_1.csv", message = "add one large dataset")
dvs_add(paths = fs::dir_ls("data/small", type = "file"), message = "add small corpus datasets")
dvs_add(glob = "data/individual/individual_*.csv", message = "add individual datasets via glob")
```

# Internals of `.dvs`

`.dvs` is hidden, so `fs::` helpers need `all = TRUE`.

```{r}
#| warning: false
setwd(new_project)

fs::dir_tree(".dvs", all = TRUE, regexp = "\\.git", invert = TRUE)
```

```{r}
#| warning: false
setwd(new_project)

fs::dir_info(".dvs", recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::select(size, path, type)
```

```{r}
#| warning: false
setwd(new_project)

fs::dir_info(".dvs", recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::mutate(subdir = path |> fs::path_dir() |> fs::path_rel(".dvs")) |>
  dplyr::reframe(n_files = dplyr::n(), size = sum(size), .by = subdir)
```

# Meta files

Each tracked data file has a `.dvs` meta file under `.dvs/`, mirroring the data tree.

```{r}
#| warning: false
setwd(new_project)

meta_files <- fs::dir_ls(".dvs", recurse = TRUE, all = TRUE, glob = "*.dvs", type = "file")
meta_files <- meta_files[!grepl("\\.git", meta_files)]
meta_files
```

Contents of the meta file for `data/large/large_1.csv`:

```{r}
#| warning: false
setwd(new_project)

large_meta <- grep("large_1", meta_files, value = TRUE)
large_meta

cat(readLines(large_meta, warn = FALSE), sep = "\n")
```
# Storage internals

```{r}
fs::dir_tree(storage, all = TRUE, regexp = "\\.git", invert = TRUE)
```


# Audit log

The audit log lives in the storage directory as `audit.log.jsonl`.

```{r}
#| warning: false

audit <- fs::path(storage, "audit.log.jsonl")
audit

cat(readLines(audit), sep = "\n")
```

# Storage directory

```{r}
fs::dir_tree(storage, all = TRUE, regexp = "\\.git", invert = TRUE)
```

```{r}
fs::dir_info(storage, recurse = TRUE, all = TRUE, type = "file") |>
  dplyr::filter(!grepl("\\.git", path)) |>
  dplyr::mutate(subdir = path |> fs::path_dir() |> fs::path_rel(storage)) |>
  dplyr::reframe(n_files = dplyr::n(), size = sum(size), .by = subdir)
```

# Hashed blob is plain text

Pick up the blake3 hash from the meta file.

```{r}
#| warning: false
setwd(new_project)

meta <- jsonlite::fromJSON(large_meta)
meta

hash <- meta$hashes$blake3
hash
```

Blobs are split into `<first-2>/<rest>`:

```{r}
#| warning: false

large_blob <- fs::path(storage, substr(hash, 1, 2), substr(hash, 3, nchar(hash)))
large_blob
fs::file_exists(large_blob)
```

Initialized with `compression = "none"`, the stored object is the original CSV, addressable by hash. Readable as-is.

```{r}
#| warning: false

readLines(large_blob, n = 5)
```

```{r}
readLines(
  n = 5, fs::path(new_project, "data", "large", "large_1.csv")
)
```



# `.dvs/` mirrors `data/`

Meta files live under `.dvs/` at the same relative paths as the data files they track.

```{r}
#| warning: false
setwd(new_project)

fs::dir_tree("data", regexp = "\\.git", invert = TRUE)
```

```{r}
#| warning: false
setwd(new_project)

fs::dir_tree(".dvs", all = TRUE, regexp = "\\.git", invert = TRUE)
```

# Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage, recursive = TRUE)
```

---

**Next up**: [Lifecycle](lifecycle.html) — modifying tracked files, `unsynced` status, `dry_run` previews, and what happens when a storage blob goes missing.
