---
title: "Introduction to dvs: data version control system"
format: html
execute:
  freeze: auto
---

# Helpers

```{r}
source(here::here("vignette/mkdatasetfiles.R"))
```


See `vignette/random_files.qmd` for a demonstration of this function. 

# Setup

```{r}
library(dvs)
library(fs)
library(here)
```


```{r}
storage <- tempfile(fileext = "_storage")
new_project <- tempfile(fileext = "_project")
dir.create(storage)
dir.create(new_project)
```


Let us add 1x25MB file (large), 3x1MB (small) and 50x3MB files (individual)

```{r}
#| warning: false
setwd(new_project)

mkdatasetfiles(n_files = 1, size_mb = 25, prefix = "large_", dir = "data/large")
mkdatasetfiles(n_files = 3, size_mb = 1,  prefix = "small_", dir = "data/small")
mkdatasetfiles(n_files = 50, size_mb = 3,  prefix = "individual_", dir = "data/individual") # TODO: increase chunk size
```


```{r}
#| warning: false
setwd(new_project)

fs::dir_tree(
  "data"
)
```


```{r}
#| warning: false
setwd(new_project)

fs::dir_info(
  "data",recurse = TRUE, type = "directory",
)[, c("size", "path", "type")]
```

False sizes unfortunately;

```{r}
#| warning: false
setwd(new_project)

fs::dir_info(
  "data",recurse = TRUE, type = "file",
)[, c("size", "path", "type")] 
```


```{r}
#| warning: false
setwd(new_project)


fs::dir_info(
  "data",recurse = TRUE, type = "file",
) |> 
  dplyr::mutate(data_dir = path |> fs::path_dir() |> fs::path_rel("data")) |> 
  dplyr::select(path, type, size, data_dir) |>
  dplyr::reframe(size = sum(size), .by = c(data_dir))
```



Initialize a dvs repository

```{r}
#| warning: false
setwd(new_project)

dvs_init(storage)
```


Add the new files in the project to the dvs repository

```{r}
#| warning: false
setwd(new_project)

dvs_add("data/large/large_1.csv", message = "add one large file")
dvs_add(c("data/small/small_1.csv", "data/small/small_2.csv", "data/small/small_1.csv") , message = "add 3 small corpus datasets")
# dvs_add("data/individual", glob = "individiual_*.csv", message = "add all individual datasets")
dvs_add(glob = "data/individual/individual_*.csv", message = "add all individual datasets")

```



```{r}
#| warning: false
setwd(new_project)

dvs_add("data/many_files", glob = "*.csv", message = "added individual datasets")
```


Status

```{r}
#| warning: false
setwd(new_project)

dvs_status()
```


Let us accidentally delete a few individual files:
```{r}
#| warning: false
setwd(new_project)

unlink("data/many_files/individual_4.csv")
unlink("data/many_files/individual_8.csv")
unlink("data/many_files/individual_15.csv")
unlink("data/many_files/individual_16.csv")
unlink("data/many_files/individual_23.csv")
```


First, what is the status?

```{r}
#| warning: false
setwd(new_project)

dvs_status(status = "absent")
```


TODO: demonstrate the manual filtering with `dplyr`

```{r}

#| warning: false
setwd(new_project)

dplyr::filter(dvs_status(), status == "absent")


```



Let us restore them:

```{r}
#| warning: false
setwd(new_project)

dvs_get("data/many_files", glob = "individual_*.csv")
```





## Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage, recursive = TRUE)
```

