---
title: "Introduction to dvs"
subtitle: "a data version control system"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

# Setup

```{r}
library(dvs)
library(fs)
library(here)
```


# Helpers

```{r}
source(here::here("R/mkdatasetfiles.R"))
```


See `vignette/random_files.qmd` for a demonstration of this function. 


```{r}
storage <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())

dir.create(storage)
dir.create(new_project)
```


Let us add 1x25MB file (large), 3x1MB (small) and 50x3MB files (individual)

```{r}
#| warning: false
setwd(new_project)

mkdatasetfiles(n_files = 1,  size_mb = 25, prefix = "large_",      dir = "data/large",      show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) -> large_file
mkdatasetfiles(n_files = 3,  size_mb = 1,  prefix = "small_",      dir = "data/small",      show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) -> small_files
mkdatasetfiles(n_files = 50, size_mb = 3,  prefix = "individual_", dir = "data/individual", show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) -> individual_files
```


```{r}
#| warning: false
setwd(new_project)

fs::dir_tree(
  "data"
)
```


```{r}
#| warning: false
setwd(new_project)

fs::dir_info(
  "data",recurse = TRUE, type = "directory",
)[, c("size", "path", "type")]
```

False sizes unfortunately;

```{r}
#| warning: false
setwd(new_project)

fs::dir_info(
  "data",recurse = TRUE, type = "file",
)[, c("size", "path", "type")] 
```


```{r}
#| warning: false
setwd(new_project)


fs::dir_info(
  "data",recurse = TRUE, type = "file",
) |> 
  dplyr::mutate(data_dir = path |> fs::path_dir() |> fs::path_rel("data")) |> 
  dplyr::select(path, type, size, data_dir) |>
  dplyr::reframe(size = sum(size), .by = c(data_dir))
```

Initialize a dvs repository

```{r}
#| warning: false
setwd(new_project)

dvs_init(storage)
```

```{r}
cat(
  readLines(fs::path(new_project, "dvs.toml"))
)
```


Add the new files in the project to the dvs repository

```{r}
#| warning: false
setwd(new_project)

# single file
dvs_add("data/large/large_1.csv", message = "add one large dataset")

# whole folder
dvs_add(paths = fs::dir_ls("data/small", type = "file"), message = "add small corpus datasets")

# glob
# dvs_add("data/individual", glob = "individiual_*.csv", message = "add individual datasets via glob")
dvs_add(glob = "data/individual/individual_*.csv", message = "add individual datasets via glob")
```


Status

```{r}
#| warning: false
setwd(new_project)

dvs_status()
```


Let us accidentally delete a few individual files:
```{r}
#| warning: false
setwd(new_project)

unlink("data/individual/individual_04.csv")
unlink("data/individual/individual_08.csv")
unlink("data/individual/individual_15.csv")
unlink("data/individual/individual_16.csv")
unlink("data/individual/individual_23.csv")
```


First, what is the status?

```{r}
#| warning: false
setwd(new_project)

dvs_status(status = "absent")
```


Or filter manually with `dplyr`:

```{r}
#| warning: false
setwd(new_project)

dplyr::filter(dvs_status(), status == "absent")
```



Let us restore them:

```{r}
#| warning: false
setwd(new_project)

dvs_get(glob = "data/individual/individual_*.csv")
```


## Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage, recursive = TRUE)
```

---

This was the happy path — init, add, status, get. Everything worked as expected.

**Next up**: [CLI](intro-cli.html) — the same operations driven from the terminal via the `dvs` binary. After that, [Internals](intro-internals.html) digs into the `.dvs/` folder layout, meta files, and content-addressed blob storage.

