---
title: "dvs lifecycle"
subtitle: "modify, re-add, dry-run, and storage integrity failures"
format: html
execute:
  freeze: auto
---

# Setup

```{r}
library(dvs)
library(fs)
library(here)
```


```{r}
source(here::here("R/mkdatasetfiles.R"))
```

```{r}
storage     <- tempfile(fileext = "_storage", tmpdir = here::here())
new_project <- tempfile(fileext = "_project", tmpdir = here::here())
dir.create(storage)
dir.create(new_project)
```

```{r}
#| warning: false
setwd(new_project)
mkdatasetfiles(n_files = 1, size_mb = 25, prefix = "large_", dir = "data/large")
mkdatasetfiles(n_files = 3, size_mb = 1,  prefix = "small_", dir = "data/small")
```

```{r}
#| warning: false
setwd(new_project)
dvs_init(storage)
invisible(dvs_add("data/large/large_1.csv",
                  message = "initial add of large file"))
invisible(dvs_add(paths = fs::dir_ls("data/small", type = "file"),
                  message = "initial add of small files"))
```

All files are `current`:

```{r}
#| warning: false
setwd(new_project)
dvs_status()
```

# Modify a tracked file

Append a row to `large_1.csv` — the file now differs from the hash dvs recorded:

```{r}
#| warning: false
setwd(new_project)
cat("99,0,0,0,0,0\n", file = "data/large/large_1.csv", append = TRUE)
```

`dvs_status()` reports the hash mismatch as `unsynced`:

```{r}
#| warning: false
setwd(new_project)
dvs_status(status = "unsynced")
```

Re-add with a new message. dvs writes a new blob and overwrites the meta file:

```{r}
#| warning: false
setwd(new_project)
dvs_add("data/large/large_1.csv", message = "add row 99 — corrected dataset")
```

Back to `current`:

```{r}
#| warning: false
setwd(new_project)
dvs_status(status = "current")
```

# `dry_run`

`dry_run = TRUE` reports what *would* happen without writing anything.

Modify the file again:

```{r}
#| warning: false
setwd(new_project)
cat("100,0,0,0,0,0\n", file = "data/large/large_1.csv", append = TRUE)
```

Preview `dvs_add` without committing:

```{r}
#| warning: false
setwd(new_project)
dvs_add("data/large/large_1.csv", message = "dry run preview", dry_run = TRUE)
```

Status is still `unsynced` — nothing was written:

```{r}
#| warning: false
setwd(new_project)
dvs_status(status = "unsynced")
```

Delete a file, then preview `dvs_get` without restoring:

```{r}
#| warning: false
setwd(new_project)
unlink("data/small/small_1.csv")
dvs_get("data/small/small_1.csv", dry_run = TRUE)
```

Status still shows `absent`:

```{r}
#| warning: false
setwd(new_project)
dvs_status(status = "absent")
```

Restore for real:

```{r}
#| warning: false
setwd(new_project)
dvs_get("data/small/small_1.csv")
dvs_status()
```

# Integrity failure: missing storage blob

dvs blobs are content-addressed — the hash in the meta file is the key. If a
blob is missing from storage (e.g. accidental deletion), `dvs_get` returns an
`error` column describing the failure rather than restoring the file.

Locate and delete the blob for `small_2.csv`:

```{r}
#| warning: false
setwd(new_project)
meta_path <- grep(
  "small_2",
  fs::dir_ls(".dvs", recurse = TRUE, all = TRUE, glob = "*.dvs", type = "file"),
  value = TRUE
)
meta  <- jsonlite::fromJSON(meta_path)
hash  <- meta$hashes$blake3
blob  <- fs::path(storage, substr(hash, 1, 2), substr(hash, 3, nchar(hash)))

cat("blob:", as.character(blob), "\n")
cat("exists:", fs::file_exists(blob), "\n")
fs::file_delete(blob)
cat("exists after delete:", fs::file_exists(blob), "\n")
```

Remove the local copy so a restore is needed:

```{r}
#| warning: false
setwd(new_project)
unlink("data/small/small_2.csv")
dvs_status(status = "absent")
```

`dvs_get` returns normally but with a non-empty `error` column — the blob is
gone so the file cannot be restored:

```{r}
#| warning: false
setwd(new_project)
dvs_get("data/small/small_2.csv")
```

# Untracking files

There is **no `dvs_remove` in the current release**. To stop tracking a file,
delete its meta file under `.dvs/` manually:

```r
# dvs will no longer track large_1.csv after its meta file is removed
fs::file_delete(".dvs/data/large/large_1.csv.dvs")
```

After the meta file is gone `dvs_status()` no longer reports on that file.

# Cleanup

```{r}
unlink(new_project, recursive = TRUE)
unlink(storage,     recursive = TRUE)
```

---

**Next up**: [Collaboration](collab.html) — two projects sharing one storage directory; what to commit to git and what to gitignore.
