---
title: "set_dvs_log_level()"
subtitle: "Route core log output to the R console"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

`set_dvs_log_level()` controls which log messages from the dvs core are routed
to R's console. `error` and `warn` go to stderr; `info`, `debug`, and `trace`
go to stdout. The default at package load is `off`.

```r
set_dvs_log_level(level = c("off", "error", "warn", "info", "debug", "trace"))
```

## Parameters

| Name | Type | Default | Behavior |
|---|---|---|---|
| `level` | character(1) | `"off"` | One of `off`, `error`, `warn`, `info`, `debug`, `trace`. Validated with `match.arg()`. |

Called for its side effect; returns `NULL` invisibly.

## Setup

```{r}
options(width = 1000)
library(dvs)
proj  <- tempfile("project_")
store <- tempfile("storage_")
dir.create(proj)
dir.create(store)
write.csv(mtcars, file.path(proj, "cars.csv"))
setwd(proj)
dvs_init(store)
```

## Enable logging around an operation

Set the level to `"info"` before an add, then reset to `"off"`.

```{r}
setwd(proj)
set_dvs_log_level("info")
dvs_add("cars.csv")
set_dvs_log_level("off")
```

## Restore the default

`"off"` is the load-time default and silences core log output.

```{r}
set_dvs_log_level("off")
```

::: {.callout-note}
The R package does not consult the `RUST_LOG` environment variable;
`set_dvs_log_level()` is the only way to change the level. `RUST_LOG` applies to
the CLI binary only.
:::

## See also

- [set_dvs_threads()](r-threads.html), [dvs_version()](r-version.html)
- [Error reference](error.html)
