---
title: "dvs_status()"
subtitle: "Report the sync status of tracked files"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

`dvs_status()` reports each tracked file as `current` (on disk and matching the
stored hash), `absent` (tracked but missing on disk), or `unsynced` (on disk but
not matching the stored hash). It returns a tibble that always includes the full
metadata columns.

```r
dvs_status(
  paths = character(0),
  recursive = NULL,
  status = c("current", "absent", "unsynced")
)
```

## Parameters

| Name | Type | Default | Behavior |
|---|---|---|---|
| `paths` | character | `character(0)` (whole project) | Files or directories to report. |
| `recursive` | logical(1) | `NULL` (treated as `FALSE`) | Include files in subdirectories of given directories. |
| `status` | character | all three | Statuses to include. Validated with `match.arg(several.ok = TRUE)`. |

## Setup

Add three files, then delete one and edit another so all three states appear.

```{r}
options(width = 1000)
library(dvs)
proj  <- tempfile("project_")
store <- tempfile("storage_")
dir.create(proj)
dir.create(store)
dir.create(file.path(proj, "data", "sub"), recursive = TRUE)
write.csv(mtcars[1:8, ],  file.path(proj, "data", "f1.csv"))
write.csv(mtcars[9:16, ], file.path(proj, "data", "f2.csv"))
write.csv(iris[1:10, ],   file.path(proj, "data", "sub", "f3.csv"), row.names = FALSE)
setwd(proj)
dvs_init(store)
dvs_add(c("data/f1.csv", "data/f2.csv", "data/sub/f3.csv"), message = "initial add")
file.remove("data/f2.csv")                                   # -> absent
write.csv(mtcars[1:9, ], "data/f1.csv")                       # -> unsynced
```

## All files

With no arguments, every tracked file is reported, with all metadata columns.

```{r}
setwd(proj)
dvs_status()
```

## `paths`

Scope the report to specific files or directories.

```{r}
setwd(proj)
dvs_status("data/sub/f3.csv")
```

## `recursive`

Without `recursive`, a directory argument reports only files directly in it.
With `recursive = TRUE`, files in subdirectories are included too.

```{r}
setwd(proj)
dvs_status("data")
```

```{r}
setwd(proj)
dvs_status("data", recursive = TRUE)
```

## `status`

Pass a single status, or several via the `match.arg(several.ok = TRUE)`
mechanism.

```{r}
setwd(proj)
dvs_status(status = "absent")
```

```{r}
setwd(proj)
dvs_status(status = c("absent", "unsynced"))
```

## Selecting columns

`dvs_status()` always returns the full metadata, so there is no toggle for it.
Select the columns you want with `[]` or dplyr.

```{r}
setwd(proj)
dvs_status()[, c("path", "status", "size")]
```

```{r}
setwd(proj)
suppressMessages(library(dplyr))
dvs_status() |> select(path, status, message, add_time)
```

## Return value

A tibble with one row per file:

| Column | Type | Description |
|---|---|---|
| `path` | character | File path. |
| `status` | character | `current`, `absent`, or `unsynced`. |
| `hash` | character | blake3 content hash. |
| `size` | `dvs_bytes` | Stored size. |
| `created_by` | character | User who added the file. |
| `compression` | character | `zstd` or `none`, recorded per file. |
| `message` | character | Message from `dvs_add()`, or `NA`. |
| `add_time` | POSIXct | When the file was added. |

Rows that fail carry an `error` column. The `size` column is a
[dvs_bytes](r-bytes.html) value.

## Differences from the CLI

The CLI command is [dvs status](cli-status.html). It shows a compact `path`,
`status`, `size` table by default and only adds the metadata columns with
`--with-metadata`; `dvs_status()` always returns them. Its filters are
independent `--current` / `--absent` / `--unsynced` flags rather than a single
`status` vector, and it adds `--threads` and `--json`.

## See also

- [dvs_init()](r-init.html), [dvs_add()](r-add.html), [dvs_get()](r-get.html)
- [Storage and meta files](intro-internals.html)
