---
title: "Random Files"
format:
  html:
    keep-md: true
execute:
  freeze: auto
---

## Overview

`mkdatasetfiles()` is a test-data generator used throughout these vignettes. It is not part of
dvs. It produces a chosen number of CSV files at a chosen size, so you can exercise dvs add,
push, fetch, hash, and compression at any scale, from a single 1 MB file to multi-gigabyte
batches of hundreds of files.

Key controls:

- `n_files`: how many files to generate
- `size_mb`: target on-disk size for each file (bytes are padded to this target)
- `dir`: output directory (created automatically)
- `prefix`: filename prefix; files are numbered `<prefix>1.csv`, `<prefix>2.csv`, and so on
- `dataset`: base R data frame whose columns seed the content (default: `ChickWeight`)
- `extra_cols`: number of random-integer columns appended to each row (increases width, and
  therefore file size, independently of row count)

Each file gets its own `set.seed(i)`, so the files in a batch are distinct even though they are
the same size.

The helper is sourced inline at the top of any vignette that needs test files; nothing is
installed into the R session permanently.

---

The `mkdatasetfiles()` helper lives in `R/mkdatasetfiles.R`:

```{r}
#| echo: false
#| results: asis
cat("```r\n",
    paste(readLines(here::here("R/mkdatasetfiles.R")), collapse = "\n"),
    "\n```\n",
    sep = "")
```

Load it:

```{r}
source(here::here("R/mkdatasetfiles.R"))
```

1 small file


```{r}
mkdatasetfiles(n_files = 1,
              dir = here::here("tmp_random_files"),
              prefix = "small_",
              size_mb = 1,
              chunk_rows = 200L,
              dataset = Theoph, extra_cols = 25,
              show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) ->
  small_file
```


view the file

```{r}
read.csv(small_file, nrows = 5)
```
File size

```{r}
fs::file_info(small_file)[,c("size", "type", "path")]
```

That's a 1MB. Good!

1 large file

```{r}
mkdatasetfiles(n_files = 1,
              chunk_rows = 200000L,
              dir = here::here("tmp_random_files"),
              size_mb = 25,
              prefix = "large_",
              dataset = Theoph, extra_cols = 25,
              show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) ->
  large_file

```


```{r}
fs::file_info(large_file)[,c("size", "type", "path")]
```

```{r}
large_file_data <- read.csv(large_file)

duplicated(large_file_data) |> table(useNA = "always") |> print() |> summary()

```

2 individual files

```{r}
mkdatasetfiles(n_files = 2, dir = here::here("tmp_random_files"), size_mb = 10, prefix = "individual_", dataset = Theoph, extra_cols = 10,
              show_progress = !nzchar(Sys.getenv("QUARTO_DOCUMENT_PATH"))) ->
  individual_files
```


file info to confirm

```{r}
fs::file_info(individual_files)[,c("size", "type", "path")]

```

Are they identical?

```{r}
individual_datasets <- lapply(individual_files, \(path) read.csv(path))

all.equal(individual_datasets[[1]], individual_datasets[[2]]) |>
  isTRUE()

```

The generated individual files are not duplicates. Good.

---

Return to the [full index](index.html) for an overview of all dvs demo vignettes.

