---
title: "Setup a dvs demo project"
format: html
execute:
  freeze: auto
  eval: false
---

Follow these steps to bootstrap a new project that uses the `dvs` R package from source.

## R environment

```{r}
dir.create("demo_dvs2_rpkg")
setwd("demo_dvs2_rpkg")
usethis::create_project(".")
```

```
✔ Setting active project to ".../demo_dvs2_rpkg".
✔ Creating 'R/'.
✔ Writing a sentinel file '.here'.
☐ Build robust paths within your project via `here::here()`.
ℹ Learn more at <https://here.r-lib.org>.
✔ Setting active project to "<no active project>".
```

Restart R (Session -> Restart R), then initialize rv and add packages:

```{bash}
rv init
rv add dvs --git https://github.com/a2-ai/dvs2 --directory dvs-rpkg --branch main
rv add usethis quarto here stringr stringi dplyr
```

Optionally pin a registry: browse [prism.a2-ai.cloud](https://prism.a2-ai.cloud/) and edit `rproject.toml`:

```toml
repositories = [
  {alias = "PRISM", url = "https://prism.a2-ai.cloud/rpkgs/nimbus/2026.03/"},
]
```

## dvs source and build tool

Clone the dvs source and install the cargo build helper — this must happen before `rv sync` so the local clone is available when dvs builds from source:

```{bash}
gh repo clone A2-ai/dvs2 .dvs -- --single-branch --depth=1
echo '*' > .dvs/.gitignore

cargo install --locked --force --git https://github.com/A2-ai/miniextendr cargo-revendor
```

```
Cloning into '.dvs'...
```

```{r}
.rv$sync()
```

```
rv project successfully initialized at .

From local cache (2):
  + cli       3.6.6  binary  https://cloud.r-project.org/        1ms
  + dvs  0.3.0.9000  source  https://github.com/a2-ai/dvs2     54.1s

sync completed in 60.9s (2 installed, 0 removed)

... (usethis + deps install)
sync completed in 2.0s (30 installed, 0 removed)

... (quarto + deps install)
sync completed in 2.2s (32 installed, 0 removed)
```

## Quarto scaffolding

```{r}
usethis::create_quarto_project(".")
```

Work inside `vignette/intro.qmd`.

## Updating dvs source

To pull the latest `dvs-rpkg` source locally later:

```{bash}
cd .dvs && git pull --rebase && cd -
```
